\section{Eulerian and Hamiltonian Paths}\label{sec:euler}
\begin{definition}\leavevmode
\begin{enumerate}
    \item A \textbf{cycle}, \textbf{circuit}, or \textbf{closed path} is a path $(v_1, \ldots, v_n)$ with $v_1 = v_n$. A path which is not closed is \textbf{open}.
    \item A path $\gamma = (v_1, \ldots, v_n)$ is \textbf{simple} if $|\{v_1, v_2, \ldots, v_n\}| = n$ if $\gamma$ is open or $n-1$ if $\gamma$ is closed.
\end{enumerate}
\end{definition}

\begin{exercise} Prove that any two connected graphs with the same number of vertices and degree sequences $(2, 2, \ldots, 2)$ are isomorphic.
\end{exercise}

\begin{theorem} Let $G$ be a graph with
$$\frac{(n-1)(n-2)}{2} +1$$
edges.  Then $G$ is connected.  This bound is ``sharp'', \textit{i.e.} that this bound cannot be improved for graphs in general.
\end{theorem}

\begin{theorem} A path $\gamma= (v_1, \ldots, v_n)$ is simple if and only if for all pairs of distinct $i, j$ with $1\leq i < j\leq n$, $v_i \neq v_j$ unless perhaps $i=1$ and $j=n$.
\end{theorem}

% TODO!
\begin{examples} Do it!
\end{examples}

\begin{definition} A path \textbf{traverses} an edge $e$ if it contains the pair $v_i, v_{i+1}$ and $e=(v_i, v_{i+1})$. An \textbf{Eulerian path} is a path in a graph $G$ traverses every edge in $G$ exactly once.  An \textbf{Eulerian cycle} is a cycle which is Eulerian.
\end{definition}

% TODO!
\begin{examples} Do it!
\end{examples}

\begin{theorem} A connected graph $G$ has an Eulerian cycle if and only if every vertex of $G$ is even (\textit{i.e.} has even degree).
\end{theorem}

\begin{corollary} There is an Eulerian cycle beginning at any vertex in a graph with all even vertices.
\end{corollary}

\begin{corollary} A connected graph has an open Eulerian path if and only if every vertex is even except for exactly two.  Such a path must begin at one of the two odd vertices.
\end{corollary}

\begin{definition} An open \textbf{Hamiltonian path} is one which visits every vertex exactly once. A \textbf{Hamiltonian cycle} $\gamma$ is a cycle in which every vertex $G$ is visited exactly once, apart from the first and last vertex of $\gamma$.
\end{definition}
% TODO!
\begin{examples} Do it!
\end{examples}

\begin{lemma} If the sum of the degrees of every pair of vertices of a graph $G$ is at least $v-1$, then
\begin{itemize}
    \item every pair of vertices are either adjacent to each other or to a common third vertex, and
    \item $G$ is connected.
\end{itemize}
\end{lemma}

\begin{theorem}\leavevmode
\begin{enumerate}
    \item If the sum of the degrees of every pair of vertices of $G$ is at least $v-1$, then $G$ has an open Hamiltonian path
    \item If the sum of the degrees of every pair of vertices of $G$ is at least $v$, then $G$ has an closed Hamiltonian path
\end{enumerate}
\end{theorem}

\begin{remark} It is an important open problem in Mathematics to classify Hamiltonian cycles.  Do a little research on the current status of the problem!
\end{remark}

\begin{theorem} If $G$ is a graph such that every pair $u$ and $w$ of nonadjacent vertices, $\text{deg}(u) + \text{deg}(w) \geq v$, then $G$ is Hamiltonian.
\end{theorem}

\begin{theorem}[Dirac] If every vertex $v$ of $G$ has $\text{deg}(v) \geq \frac{1}{2}v$, then $G$ is Hamiltonian.
\end{theorem}

\begin{remark} Research the K\"{o}nigsberg Bridge Problem.  We haven't discussed it here because it requires the concept of multigraphs (\textit{i.e.} graphs with multiple edges between pairs of vertices).  State the problem, and justify the answer to the problem in terms of multigraphs.
\end{remark}
