\section{Set Theory}\label{sec:settheory}

\begin{definition} \leavevmode
\begin{enumerate}
    \item \label{def:set} A \textbf{Set} is a collection of distinct objects, none of which is the set itself.  If $a$ is an object belonging to the set $A$, we write $a \in A$, and say ``$a$ is an element of $A$''.
    \item \label{def:nullset} A set containing no elements is called the \textbf{empty set}, or the \textbf{null set}, and is written $\emptyset$ or $\{ \}$.
    \item \label{def:subset} A set $A$ is said to be a \textbf{subset} of the set $B$, written $A \subseteq B$, if every element of $A$ is also an element of $B$.
    \item \label{def:setequality} A set $A$ is said to be a \textbf{equal to} the set $B$, written $A = B$ if $A \subseteq B$ and $B \subseteq A$.
\end{enumerate}
\end{definition}

If it is possible to enumerate the elements of $A$, we do so with the following notation:
$$ A = \left\{ a, \pi, \frac{45}{36}, \text{``Massachusetts''}\right\}.$$

\begin{remark} You may find the definition of a mathematical set nebulous and confusing.  What's a ``collection''?  What's an ``object'', and what does it mean for them to be ``distinct''?  In truth, while it is possible to formally define all of these concepts, it is typically the case that a student has an intuitive understanding of a set, and can begin from that.
\end{remark}

\begin{exercise} List all the subsets of $\{1, 2, 3\}$.
\end{exercise}

\begin{notation} (Set Builder Notation) Let $A$ be a set, and for all $x \in A$, let $p(x)$ be a proposition about $x$ which may be true or false.  Then we may build a set $S$ by taking all those elements of $A$ for which the proposition is true; such a set may be written down using \textbf{set builder notation}:
    $$S = \{x\in A\mid p(x)\},$$
and we read this as ``S is (equal to) the set of all $x$ in $A$ such that $p$ of $x$.''  One important note is that a set $A$ must exist in order to use set builder notation; as a result of this, we will use the term \textit{universe of discourse}, often denoted by $X$, to describe any reasonably conceivable objects that may be placed into a set.  You will see this appearing in the definitions ahead (see for example Definition \ref{def:setunion}).
\end{notation}

\begin{exercises} Let $\mathbb{N} = \{1, 2, 3, \ldots\}$ denote the set of natural numbers.\begin{enumerate}
    \item Translate the set $S=\{1, 2, 3, 4, 5\}$ into set builder notation, where $A=\mathbb{N}$.
    \item Write down, without the uses of ellipses (``$\ldots$''), notation defining the set of even natural numbers
    \item Repeat the above exercise for the set of odd natural numbers divisible by 5 (one may use ``$7\mid 14$'' to say that ``$7$ divides $14$'').
\end{enumerate}\end{exercises}

\begin{theorem}\label{thm:nullsetunique} There is only one empty set.
\end{theorem}

\begin{theorem}[transitivity of subset]\label{thm:subsettransitivity}  If $A \subseteq B$ and $B \subseteq C$, then $A \subseteq C$.
\end{theorem}

\subsection{Getting new sets from old}
\begin{definition} \label{def:setunion} Let $A$ and $B$ be sets, and let $X$ denote the universe of discourse.
\begin{enumerate}
    \item  The set $A \cup B = \{x\in X\mid x \in A \lor x\in B\}$ is called the \textbf{union} of $A$ and $B$.
    \item \label{def:setintersection} The set $A \cap B = \{x\in X\mid x \in A \land x\in B\}$ is called the \textbf{intersection} of $A$ and $B$.
    \item \label{def:setcomplement} The set $A \smallsetminus B = \{x\in A\mid x \not\in B\}$ is called the \textbf{(relative) complement} of $A$ in $B$.
\end{enumerate}
\end{definition}

\begin{theorem} For all sets $A$ and $B$, if $A \subseteq A \cap B$ then $A \cup B \subseteq B$.
\end{theorem}

\begin{theorem} For all sets $A$, $B$, $C$, and $D$, if $A\subseteq C$ and $B\subseteq D$ then $A\cup B \subseteq C\cup D$.
\end{theorem}

\begin{theorem} For all sets $A$, $B$, $C$, and $D$, if $A\subseteq C$ and $B\subseteq D$ then $A\cap B \subseteq C\cap D$.
\end{theorem}

\begin{theorem} \label{thm:reversesetcomplements}Let $A$, $B$, and $X$ be sets. If $A \subseteq B$, then $X \smallsetminus B \subseteq X \smallsetminus A$.
\end{theorem}

\begin{exercise} Write down the \textit{converse} of Theorem \ref{thm:reversesetcomplements}. It isn't quite true (why not?) unless you also add the assumption that $A\subseteq X$ and $B\subseteq X$.  Prove the converse with that additionally assumption. (The \textit{converse} of a statement $p \to q$ is $q \to p$.)
\end{exercise}

\begin{theorem} Let $A$ and $B$ be sets. Then $A\smallsetminus B = \emptyset$ if and only if $A \subseteq B$.
\end{theorem}

\begin{theorem} For sets $A$ and $B$, $(A\smallsetminus B)\cup (B\smallsetminus A) = (A\cup B) \smallsetminus (A\cap B)$.
\end{theorem}

For our purposes, a \textit{claim} is something that may or may not be true, and we need to determine whether or not it is true and justify it with a proof or counterexample.

\begin{claim} For all sets $A$, $B$, and $C$, if $A \subseteq B\cup C$ then $A\subseteq B$ or $A\subseteq C$.
\end{claim}

\subsection{Bijections and cardinality}
\begin{definition} \label{def:bijection} Let $A$ and $B$ be sets.
\begin{enumerate}
    \item Let $a\in A$ and $b\in B$.  Then the \textbf{ordered pair} of $a$ and $b$, written $(a, b)$, is pairing of the elements $a$ and $b$ into an ordered grouping.  Strictly speaking (though this intuitive definition typically suffices), one may define $(a, b)=\{\{a\}, \{a, b\}\}$.  We refer to $a$ and $b$ as \textit{elements} of $(a,b)$, even though strictly speaking they are not.
    \item A \textbf{bijection}, or a \textbf{one-to-one correspondence}, between $A$ and $B$ is a set $C$ with all of the following properties.
    \begin{itemize}
        \item Every element of $C$ consists of an ordered pair $(a,b)$ where $a \in A$ and $b \in B$.
        \item (injective) Every element of $A$ exists as the first element of exactly one element of $C$.
        \item (sujective) Every element of $B$ exists as the second element of exactly one element of $C$.
    \end{itemize}
    We say that $A$ and $B$ are \textbf{in bijection} (or sometimes \textit{bijective}) if there exists a bijection between them; this is sometimes written as $A \cong B$, but it often just written out in words.
\end{enumerate}
\end{definition}

\begin{remark} In a traditional set theory course, one uses ordered pairs to first define cartesian products, and then relations, functions, injections, surjections, domain, co-domain, range, \textit{etc.} before defining bijections. For our purposes, bijections as defined above will suffice.
\end{remark}

\begin{theorem}[Bijectivity is an equivalence relation] Let $A$, $B$, and $C$ be sets.
\begin{enumerate}
    \item (reflexivity) $A$ is in bijection with itself.
    \item (symmetry) If $A$ is in bijection with $B$, then $B$ is in bijection with $A$.
    \item (transitivity) If $A$ is in bijection with $B$, and $B$ is in bijection with $C$, then $A$ is in bijection with $C$.
\end{enumerate}\end{theorem}

\begin{remark} The fact that bijections satisfy the above three properties give it the status of being what's called an \textbf{equivalence relation}.  We will see equivalence relations again in the future when we discuss graphs.  One often considers equivalence relations to be a notion of ``sameness'': if $A$ is in bijection with $B$, then they're essentially the same object.
\end{remark}

\begin{definition} If a set $A$ is in bijection with the set $\{1, 2, 3, 4, \ldots, n\}$, then the \textbf{cardinality} of $A$ is given by $n$, written $|A| = n$, and we say that $A$ is \textbf{finite}.  If a set is in bijection with the natural numbers, then we say that it is \textbf{countably infinite}.\end{definition}

\begin{theorem} \label{thm:bijectioncardinality} If $|A|\neq |B|$, then $A$ is not in bijection with $B$. \end{theorem}

\begin{question} Is the \textit{converse} of Theorem \ref{thm:bijectioncardinality} true?  Prove or disprove.
\end{question}

\begin{theorem} Being countably infinite and finite are mutually exclusive set properties. \end{theorem}




\subsection{Exercises}
\begin{enumerate}
    \item Let $A$, $B$, and $C$ be sets.  Prove that if $A \subseteq C$ and $B \subseteq C$ then $A \cup B \subseteq C$.

    \item Given a set $A$ with $|A|=n$, how many subsets does $A$ have?  Prove your answer.

    \item Prove that the natural numbers are in bijection with the even natural numbers.

    \item Prove that the natural numbers are in bijection with the integers.

    \item Let $C$ be a bijection between the natural numbers ($\mathbb{N}$) and the integers ($\mathbb{Z}$), so that $C\subseteq\{(x,y)\mid x\in \mathbb{N}\land y\in\mathbb{Z}\}$.  Show that there exist elements $(a, b)$ and $(x,y)$ in $C$ such that $a > x$ and $b < y$.

    \item Prove that the natural numbers are in bijection with the set of ordered pairs $\{(n, a) \mid n\in \mathbb{N} \land a \in \{1,2,3\}\}$.

    \item Prove that the natural numbers are in bijection with the set of ordered pairs $\{(n, m) \mid n\in \mathbb{N} \land a \in \mathbb{N}\}$.

    \item Prove that the set of words in this sentence is not in correspondence with the set of words in the preamble to the U.S. Constitution.
\end{enumerate}

\subsection{Addendum: Bijective Functions}
This section originally appeared a single sheet, with a theorem labeled ``Theorem 1''.

\begin{definition} \label{def:bijective_function} Let $A$ and $B$ be sets.
\begin{enumerate}
    \item A set $C$ of the form $C=\{x\in X\ |\ \exists a\in A\ \exists b\in B\ x = (a,b)\}$ is called a \textbf{relation} between $A$ and $B$.
    \item  A \textbf{function} from $A$ to $B$, written $f:A\to B$, is a relation $C = C(f)$ that has the additional properties:
    \begin{enumerate}
        \item For all $a \in A$, there exists an element $b\in B$ such that $(a,b)\in C$, and
        \item For all $a \in A$ and $b_1, b_2 \in B$, if $(a,b_1)\in C$ and $(a,b_2)\in C$ then $b_1 = b_2$.
    \end{enumerate}
    \item A function is said to be \textbf{injective}, or \textbf{one-to-one}, if it has the property: for all $a_1, a_2 \in A$ and $b\in B$, if $(a_1,b)\in C$ and $(a_2,b)\in C$ then $a_1 = a_2$.
    \item A function is said to be \textbf{surjective}, or \textbf{onto}, if for all $b \in B$, there is an $a\in A$ such that $(a, b)\in C$.
    \item A \textbf{bijective function}, also called a \textbf{one-to-one correspondence}, is a function which is injective and surjective.
\end{enumerate}
\end{definition}

We typically use the notation $f(a) = b$ to denote that $(a,b)\in C(f)$.  So then the two pieces of the definition of a function $f$ become what you're likely used to:
\renewcommand{\theenumi}{\alph{enumi}}
\begin{enumerate}
    \item For every $a \in A$, there exists an element $b\in B$ such that $f(a)=b$, and
    \item For all $a \in A$ and $b_1, b_2 \in B$, if $f(a) =b_1$ and $f(a) = b_2$ then $b_1 = b_2$.
\end{enumerate}
These are colloquially ``the domain is all of $A$'' and ``$f$ passes the vertical line test,'' respectively.  The definition of injective and surjective become:
\begin{itemize}
    \item The function $f:A\to B$ is injective if for all $a_1, a_2 \in A$ and $b\in B$, if $f(a_1)=b$ and $f(a_2) = b$ then $a_1 = a_2$.
    \item The function $f:A\to B$ is surjective if for all $b \in B$, there is an $a\in A$ such that $f(a)= b$.
\end{itemize}
After you do this change of notation, you never need to discuss $C$ again, so it is dropped from the discussion.  What we'd like to do is equate the two definitions of ``bijective'', so that one has access to whichever definition makes more sense to them.

\begin{theorem} Let $A$ and $B$ be sets.
\begin{enumerate}
    \item If $C$ is a bijection between $A$ and $B$ (in the sense of Definition \ref{def:bijection}), then there is a bijective function $f:A \to B$ such that for all $a \in A$, the (unique) element $b\in B$ such that $(a, b)\in C$ satisfies $f(a) = b$.
    \item If $f:a\to b$ is a bijective function (in the sense of Definition \ref{def:bijective_function}), then the set $C=\{x\in X\ |\ \exists a\in A\ x = (a,f(a))\}$ is a bijection.
\end{enumerate}
\end{theorem}
In other words, for any sets $A$ and $B$, there is a bijection between the set of all bijections between $A$ and $B$ in the sense of Definition \ref{def:bijection} and the set of all bijective functions between $A$ and $B$ in the sense of Definition \ref{def:bijective_function}.
